---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In the installed packadge SQL we are able to filter columns by using SELECT and rows by using WHERE. For refrence we will use and example from the Lahman Baseball database. 

* Load the Lahman Packadge and also the sqldf packadge

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)

```

* Example 1: Display table containing playerID, teamID, yearID, and HR for teamID='NYA' and yearID=1927.

```{r}
query<-"select playerID,teamID,yearID,HR from Batting
Where teamID='NYA' and yearID=1927"
sqldf(query)
```

----------------------------------------

##Locating and Specifying

Similar to the execution above we further explored specifying and locating certain criteria by use of subject in question (i.e. playerID,HR,teamID etc.) by instead of just using '=' we pecified with '<','>','<=','>=', and so on. 


* Example 1: Find more than 40 HR and less than 60 SO.

```{r}
query<-"select playerID, yearID, teamID, HR, SO from Batting where HR>40 and SO<60"
sqldf(query)
```


* Example 2: Find more than 30 HR, in the 1970's, on the Phillies 'PHI'.

```{r}
query<-"select playerID, yearID, teamID, HR from Batting where HR>30 and yearID<=1979 and yearID>=1970 and teamID='PHI'"
sqldf(query)
```


* Example 3: Find the players on the NYA that had greater than or equal to 40 HR

```{r}
query<-"select playerID, yearID, teamID, HR from Batting where teamID='NYA' and HR>=40"
sqldf(query)
```

-----------------------------------

## Order By

With the use of SQL you can order rows. To do this specification I would add "order by" to my string and specify with "desc"

* Example 1: Display the players with more than 50 HR and from greatest to least. 

```{r}
query<-"select playerID, yearID, teamID, HR from Batting 
where HR>50
order by HR desc"

sqldf(query)
```

* Example 2: Specify to display players with less than 10 strike-outs with 400 or more at-bat ordered by strikeouts.

```{r}
query<-"select playerID, yearID, teamID, SO, AB 
from Batting where SO<10 and AB>=400 order by SO"
sqldf(query)
```

----------------------------------------

##Aggregation

In order to aggregate data sets together one may use the commands sum, avg, min, max to bring together under one variable. 

* Example 1: Display the sum of Babe Ruth's total HR, specify with 'sum(HR)' and grouped by playerID.
```{r}
query<-"select playerID, sum(HR) from Batting where playerID='ruthba01'
group by playerID"
sqldf(query)
```

* Example 2: Display players that have 600 or more HR for their career total, in descending order. 
```{r}
query<-"select playerID, sum(HR) from Batting 
group by playerID 
having sum(HR)>=600
order by sum(HR) desc"
sqldf(query)
```

* Example 3: Find if a players average HR is greater than 30.
```{r}
query<-"select playerID, avg(HR) from Batting 
group by playerID 
having avg(HR)>30
order by avg(HR) desc"
sqldf(query)
```

------------------------------------------------

#Joins

When one wants to join two tables together one must specify with the use of 'inner join' when specifying from. In addition one must specify which variables are shared between the two tables.

* Example 1: Join Batting table with Master table to display nameFirst and nameLast instead of playerID as 'ruthba01'.
```{r}
query<-"select nameFirst, nameLast, teamID, yearID,HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where Batting.playerID = 'ruthba01'"

sqldf(query)
```

* Example 2: Join Batting and Master to display names of players with HR greater than 50 in descending order.
```{r}
query<-"select nameFirst, nameLast, teamID, yearID, HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where HR>50
order by HR desc"

sqldf(query)
```

* Example 3: Join Batting and Teams tables as to display the full name of team instead of teamID for the player 'ruthba01'.
```{r}
query<-"select playerID, name, Batting.yearID, Batting.HR
from Batting inner join Teams
on Batting.teamID=Teams.teamID
and Batting.yearID=Teams.yearID
where playerID='ruthba01'"

sqldf(query)
```

* Example 4: Joining of three tables together(Batting, Teams, Master) and identifying 'ruthba01'. When joining three tables use '()' when first join is written. 

```{r}
query<-"select nameFirst, nameLast, name, Batting.yearID, Batting.HR 
from (Batting inner join Master 
on Batting.playerID=Master.playerID) inner join Teams 
on Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
where Batting.playerID='ruthba01'"

sqldf(query)
```


* Example 5: Join Batting and Master to display players that have a career total of more than 600 HR.
```{r}
query<-"select nameFirst, nameLast, sum(HR)
from Batting inner join Master 
on Batting.playerID = Master.playerID
group by Batting.playerID
having sum(HR)>600
order by sum(HR) DESC"

sqldf(query)
```


* Example 6: Combine tables Salaries and Master to find the average salary greater than 1 million dollars of all players with the first name 'Bob'.
```{r}
query<-"select nameFirst, nameLast, avg(salary) 
from Salaries inner join Master
on Salaries.playerID=Master.playerID
where nameFirst='Bob'
group by Salaries.playerID
having avg(salary)>1000000
order by avg(salary) desc"

sqldf(query)
```

